---
title: "WGNCA_SmartSeq_V2"
output: html_document
---

# Set environment
```{r}
library(dplyr)
library(Seurat)
library(WGCNA)
```

Data came from: https://portal.brain-map.org/atlases-and-data/rnaseq/human-multiple-cortical-areas-smart-seq 

# Limit data set to only motor cortex (MC)
M1lm #lower limbs 
M1ul #upper limbs
```{bash}
# Take the sample IDs from the cells originating from the MC
cat metadata.csv | tr "," "\t" | grep -w 'M1ul\|M1lm' | awk '{print $1}' | sort -k1,1n > SampleID_MC_srt    #6235 cells
# Limit the count matrix to the cells of the MC
cat matrix.csv | sort -t ',' -k1,1n > matrix_srt.txt

join -t ',' SampleID_MC_srt matrix_srt.txt > C
cat  matrix.csv | head -n1 > matrix_MC.csv
cat C  >> matrix_MC.csv

# Limit the meta data matrix to the cells of the MC
cat metadata.csv | sort -t ',' -k1,1n > metadata_srt.txt

join -t ',' SampleID_MC_srt metadata_srt.txt > Y
cat  metadata.csv | head -n1 > metadata_MC.csv
cat Y  >> metadata_MC.csv
```

# Load the data
```{r}
MC.data <-read.csv('~/Documents/Minor research project/snSeq_MotorCortex/snSeq_MotorCortex/data/matrix_MC.csv', row.names = 1)
MC.meta.data <- read.csv('~/Documents/Minor research project/snSeq_MotorCortex/snSeq_MotorCortex/data/metadata_MC.csv', row.names = 1)
```

# Make the MC Seurat object
```{r}
# Make rows into columns and the other way around  
MC.data1 = data.frame(t(MC.data))

# Make the Seurat object
MC <- CreateSeuratObject(counts = MC.data1,  min.cells = 10, min.genes = 10, project = "MC")

# Add meta data
MC <-  AddMetaData(object = MC, 
                       metadata = MC.meta.data)
```

# Calculate percentage mito and max UMI
```{r}
# % Mitochondrial genes
MC[["percent.mt"]] <- PercentageFeatureSet(MC, pattern = "^MT-")

##Z-score nUMI on a per sample basis
nUMI.z <- data.frame(rownames(MC@meta.data), MC@meta.data$nCount_RNA)
colnames(nUMI.z) <- c("cell",  "nUMI")
nUMI.z <- nUMI.z  %>% mutate(nUMI.z = scale(nUMI))
nUMI.z <- as.data.frame(nUMI.z)
rownames(nUMI.z) <- nUMI.z$cell
nUMI.z <- subset(nUMI.z, select = "nUMI.z")
MC <-  AddMetaData(object = MC, metadata = nUMI.z, col.name = "nUMI.z")

max.nUMI.MC <- max(MC@meta.data$nUMI)      
```

# Normalize the data
```{r}
MC <- NormalizeData(MC, normalization.method = "LogNormalize", scale.factor = max.nUMI.MC)
MC <- FindVariableFeatures(MC, selection.method = "vst", nfeatures = 2000)
```

# Scale the data 
```{r}
rownames(MC)->all.genes
ScaleData(MC, features=all.genes)->MC
```

#PCA
```{r}
MC <- RunPCA(object = MC, 
                 features = VariableFeatures(object = MC))
```

# Cluster cells
```{r}
MC <- FindNeighbors(MC, dims=1:20)
MC <- FindClusters(object = MC, resolution = 0.8)
```

# TSNE
```{r}
MC <- RunTSNE(MC,
                  reduction.use = "pca",
                  dims.use = 1:20,
                  perplexity = 100)
```


# Subset for clusters and remove genes with expression in less than 50% of cells
```{r}
# Subset the 18 clusters
clusters<-as.vector(c(1:18))
for (cluster in clusters) {
  name=as.character(paste("MC.c",cluster, sep=""))
  assign(name, subset(MC, idents = as.numeric(paste(cluster))))
}

# Extract the normalized data 
for (cluster in clusters) {
  name=as.character(paste("MC.c", cluster, ".norm", sep=""))
  y=(paste0("MC.c", cluster))
  object=mget(y)
  x=object[[1]]@assays$RNA@data
  assign(name, as.data.frame(t(as.matrix(x))))
}

# Remove genes with over 50% zero values
for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".norm"))
  object=mget(y)
  ##  Take half the number of cells 
  nrow(object[[1]])->X; as.numeric(X)->X; X/2->X
  ## Find the colums that have over half zero values
  i <- colSums(object[[1]] == 0, na.rm=TRUE) < X
  ## Remove these columns
  name<-paste0("MC.c", cluster, ".keep", sep="")
  assign(name, object[[1]][, i, drop=FALSE])
}

# Scale the data
for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".keep"))
  object=mget(y)
  name=paste0("MC.c", cluster, ".scaled", sep="")
  assign(name, as.data.frame(scale(object[[1]])))
}

```

# 1. Data input, cleaning, and processing
```{r}
# Check for genes or cells with too many missing values
for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".scaled"))
  object=mget(y)
  name=paste0("MC.c", cluster, ".gsg", sep="")
  assign(name,  goodSamplesGenes(object[[1]], verbose = 3, minFraction = 1/2))
}

# Check if they all return TRUE
MC.c1.gsg$allOK

# Cluster the cells to see if there are any obvious outliers
for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".scaled"))
  object=mget(y)
  sampleTree = hclust(dist(object[[1]]), method = "average");
  sizeGrWindow(12,9)
  par(cex = 0.6);
  par(mar = c(0,4,2,0))
  plot(sampleTree, main = paste("Sample clustering to detect outliers","cluster",cluster), sub="", xlab="", cex.lab = 1.5,
  cex.axis = 1.5, cex.main = 2)
}

```

## b. Step-by-step network construction and module detection:
https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-02-networkConstr-man.pdf

### Calculate the Soft Power Threshold
```{r}
options(stringsAsFactors = FALSE);

# Automatic construction of the gene networks and identfication of modules: 
## Choosing the soft-thresholding power: analysis of network topology. The function pickSoftThreshold  performs the analysis of network topology and aids the user in choosing a proper soft-thresholding power. The soft threshold is to reduce noice correlation and thereby creates a more centralized network (a few strong correlations, lot weak correlations). The user chooses a set of candidate powers (the function provides suitable default values), and the function returns a set of network indices that should be inspected, for example as follows:

# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))

for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".scaled"))
  object=mget(y)
  # Call the network topology analysis function
  sft=pickSoftThreshold(object[[1]], powerVector = powers, verbose = 5)
  # Plot the results:
  sizeGrWindow(9, 5)
  par(mfrow = c(1,2));
  cex1 = 0.9;
  # Scale-free topology fit index as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n", main = paste("Scale independence","cluster",cluster));
    text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
    labels=powers,cex=cex1,col="red");
    # this line corresponds to using an R^2 cut-off of h
    abline(h=0.90,col="red")
  # Mean connectivity as a function of the soft-thresholding power
    plot(sft$fitIndices[,1], sft$fitIndices[,5],
    xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
    main = paste("Mean connectivity", "cluster",cluster))
    text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
}
```

### Calculate Adjacencies and make TOM
```{r}
options(stringsAsFactors = FALSE);

# We choose the power 5, which is the lowest power for which the scale-free topology fit index curve flattens out upon reaching a high value (in this case, roughly 0.90)
softPower=5

# Calculate adjacencies, TOM, and dissTOM
 ## To minimize effects of noise and spurious associations, we transform the adjacency into Topological Overlap Matrix, and calculate the corresponding dissimilarity:
  ### TOM is a measure for how many first neighbors different nodes share. Can correct for missing of faulty connections.
for (cluster in clusters) {
  y=(paste0("MC.c",cluster ,".scaled"))
  object=mget(y)
  name=paste("adjacency.", cluster, sep="")
  assign(name, adjacency(object[[1]], power=softPower))
  object1=mget(paste("adjacency.", cluster, sep=""))
  name1=paste("TOM", cluster, sep="")
  assign(name1, TOMsimilarity(object1[[1]]))
  object2=mget(paste("TOM", cluster, sep=""))
  name2=paste("dissTOM", cluster, sep="")
  assign(name2, 1-object2[[1]])
}

```

### Hierarchical clustering to produce a hierarchical clustering tree (dendrogram) of genes.
```{r}
for (cluster in clusters) {
  object=mget(paste("dissTOM", cluster, sep=""))
  # Call the hierarchical clustering function
  assign(paste("geneTree.",cluster,sep=""), hclust(as.dist(object[[1]]), method = "average"));
  # Plot the resulting clustering tree (dendrogram)
  object=mget(paste("geneTree.",cluster, sep=""))
  sizeGrWindow(12,9)
  plot(object[[1]], xlab="", sub="", main = paste("Gene clustering on TOM-based dissimilarity","cluster",cluster),
  labels = FALSE, hang = 0.04)
}

  # Each vertical line corresponds to a gene. Branches of the dendogram group together densely interconnected, highly co-expressed genes. Module identification amounts to the identification of individual branches (”cutting the branches off the dendrogram”). There are several methods for branch cutting; our standard method is the Dynamic Tree Cut from the package dynamicTreeCut. The next snippet of code illustrates its use
# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 20;

for (cluster in clusters) {
  object=mget(paste("geneTree.",cluster ,sep=""))
  object1=mget(paste("dissTOM", cluster, sep=""))
  name=paste("dynamicMods.",cluster,sep="")
  assign(name, cutreeDynamic(dendro = object[[1]], distM = object1[[1]], deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = minModuleSize))
  print(paste("This is cluster", cluster))
  print(table(mget(name)))
  # Label 0 is reserved for unassigned genes. 
  # Plot the module assignment under the gene dendrogram:
  # Convert numeric lables into colors
  name1=paste("dynamicColors.", cluster, sep="")
  assign(name1, labels2colors(mget(name)))
  object2=mget(name1)
  print(table(object2[[1]]))
  # Plot the dendrogram and colors underneath
  sizeGrWindow(8,6)
  plotDendroAndColors(object[[1]], object2[[1]], "Dynamic Tree Cut", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, main = paste("Gene dendrogram and module colors", "cluster", cluster))
}  
```

### Merging of modules whose expression profiles are very similar.
```{r}
  # The Dynamic Tree Cut may identify modules whose expression profiles are very similar. It may be prudent to merge such modules since their genes are highly co-expressed. To quantify co-expression similarity of entire modules, we calculate their eigengenes (sort of like PCA) and cluster them on their correlation:
for (cluster in clusters) {
  object=mget(paste0("MC.c",cluster,".scaled"))
  dynamicColor=mget(paste0("dynamicColors.",cluster))
  # Calculate eigengenes
  X=moduleEigengenes(object[[1]], dynamicColor[[1]])
  assign(paste0("MEs.", cluster),-X$eigengenes)
  object1=mget(paste0("MEs.", cluster))
  # Calculate dissimilarity of module eigengenes
  Y=1-cor(object1[[1]])
  # Cluster module eigengenes
  assign(paste0("METree", cluster), hclust(as.dist(Y), method = "average"))
  object2=mget(paste0("METree", cluster))
  sizeGrWindow(7, 6)
  plot(object2[[1]], main = paste("Clustering of module eigengenes","cluster",cluster),
  xlab = "", sub = "")
  MEDissThres = 0.25
  # Plot the cut line into the dendrogram
  abline(h=MEDissThres, col = "red")
}

# Merge the close modules
MEDissThres = 0.25

for (cluster in clusters) {
 object=mget(paste0("MC.c",cluster,".scaled")) 
 object1=mget(paste0("geneTree.",cluster))
 dynamicColor=mget(paste0("dynamicColors.",cluster))
 X <-mergeCloseModules(object[[1]], dynamicColor[[1]], cutHeight = MEDissThres, verbose = 3)
 assign(paste0("moduleColors.", cluster), X$colors)
 assign(paste0("MEs.", cluster), X$newMEs)
 sizeGrWindow(12, 9)
 plotDendroAndColors(object1[[1]], cbind(dynamicColor[[1]], X$colors), c(paste("Dynamic Tree Cut", "cluster", cluster), paste("Merged dynamic", "cluster", cluster)), dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05)
} 

# # Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
for (cluster in clusters) {
  object=mget(paste0("moduleColors.", cluster))
  assign(paste0("moduleLabels.", cluster), match(object[[1]],colorOrder)-1)
} 

```

# Write text file for every module for every cluster
```{r}
for (cluster in clusters) {
  object=mget(paste0("MC.c",cluster,".scaled"))
  object=object[[1]]
  moduleColor=mget(paste0("moduleColors.",cluster))
  moduleColor=moduleColor[[1]]
  modules=moduleColor%>%unique
  print(moduleColor)
  n=length(modules)
  for (i in 1:n) {
    name=paste0("Cluster",cluster,"_",modules[i],".csv")
    write.csv(names(object[moduleColor==modules[i]]), file=name)
  }
}
```

# Number of modules and genes within modules per cluster
```{r}

for (cluster in clusters) {
   moduleColors=mget(paste0("moduleColors.",cluster))
   moduleColors=moduleColors[[1]]
   print(paste("This is cluster", cluster))
   print(table(moduleColors))
} 
    
```


# 5. Network visualization using WGCNA functions
https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-05-Visualization.pdf 

## b. Exporting to Cytoscape
```{r}
# Recalculate topological overlap if needed
for (cluster in clusters) {
   object=mget(paste0("MC.c",cluster,".scaled")) 
   TOM = TOMsimilarityFromExpr(object[[1]], power = 5)
   probes=names(object[[1]])
   moduleColors=mget(paste0("moduleColors.",cluster))
   moduleColors=moduleColors[[1]]
   modules=moduleColors%>%unique
   modules=str_remove(modules, 'turquoise')
   inModule = is.finite(match(moduleColors, modules));
   modGenes = probes[inModule];
   modTOM = TOM[inModule, inModule];
   dimnames(modTOM) = list(modGenes, modGenes)
   cyt = exportNetworkToCytoscape(modTOM, edgeFile = paste("Cluster",cluster,"-CytoscapeInput-edges-", paste(modules, collapse="-"), ".txt", sep=""), nodeFile = paste("Cluster",cluster,"-CytoscapeInput-nodes-", paste(modules, collapse="-"), ".txt", sep=""), weighted = TRUE, threshold = 0.02, nodeNames = modGenes, nodeAttr = moduleColors[inModule]);
}   

# Open with Cytoscape: import the edges file by using the function "import network from file" and thereby specifying which columns are target and which are source nodes.
```

```{r}
# FUS-containing clusters
## CLuster 1 | brown
## Cluster 18 | black
## Cluster 12 | green

# Select modules
modules = c("green");
# Select module probes
probes = names(MC.c12.scaled)
inModule = is.finite(match(moduleColors.12, modules));
modGenes = probes[inModule];

# Select the corresponding Topological Overlap
modTOM = TOM12[inModule, inModule];
dimnames(modTOM) = list(modGenes, modGenes)

# Export the network into edge and node list files Cytoscape can read
cyt = exportNetworkToCytoscape(modTOM,
edgeFile = paste("C12_CytoscapeInput-edges-", paste(modules, collapse="-"), ".txt", sep=""),
nodeFile = paste("C12_CytoscapeInput-nodes-", paste(modules, collapse="-"), ".txt", sep=""),
weighted = TRUE,
threshold = 0.02,
nodeNames = modGenes,
altNodeNames = modGenes,
nodeAttr = moduleColors[inModule]);
```

